Supplementary Information

### 2021-08-31 doc/manuscript/ForestGEO_treerings_SI.Rmd
various minor/ formatting fixes to SI
- table footnotes
- resizing plots to fit
- knit in manually fixed climwin plots (issue #145 )

# climwin plots
do.call(file.remove, list(list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-tree-rings/doc/manuscript/tables_figures/SI_figures/climwin_plots_combined/", full.names = TRUE))) # first remove old files
flist <- list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-tree-rings/results/climwin_plots_combined/", full.names = TRUE)
file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-tree-rings/doc/manuscript/tables_figures/SI_figures/climwin_plots_combined/")

@@

@@

$$DBH_Y  = DBH_{m} - 2*\left[r_{bark,m} - r_{bark,Y} + \sum_{y=Y}^{m} RW_{y} \right]$$
Here, $r_{bark}$ denotes bark thickness, which was estimated on a species-specific basis from allometries (Tables `r species_table`, `r bark_allometries`).
If $DBH_m$ was taken before coring, RW values between $y=m$ and $y=Y$ were added instead of subtracted.
If $DBH_m$ was taken $\le$ 10 years after the last year of measurement in the core, we used the equation above, assuming that the $n$ missing measurements between the year of the last RW record in the core and year $m$ were equal to the last $n$ records in the core.

@@

$$DBH_Y  = 2*\left[r_{bark,Y} + \sum_{y=0}^{Y} RW_{y} \right]$$
Again, $r_{bark}$ was estimated on a species-specific basis from allometries (Tables `r species_table`, `r bark_allometries`).

@@

Regardless of the DBH reconstruction approach (outside-in or inside-out), accurate reconstructions require accounting for $r_{bark}$, particularly for species that can attain very thick bark (e.g., *Pseudotsuga menziesii*, *Pinus ponderosa* P. Lawson & C. Lawson).
For the majority of species, including all thick-barked species, we assigned allometries relating $r_{bark}$ to DBH, which were constructed for the species of interest or a related species (Tables `r species_table`, `r bark_allometries`).
When raw data on bark thickness data were available, we generated species-specific lognormal scaling relationships with the intercept forced to zero (Table `r bark_allometries`). 
When raw bark thickness data were not available, we applied published bark allometries from other sources (Table `r bark_allometries`).

@@

*Qualitative comparison*

@@

*Quantitative comparison*

@@

**Bark allometry field indicates the species and site sampled to construct the bark allometry. When neither raw data nor an allometric equation for the study species was available, we selected the most appropriate equation that could be located for similar species. Equations are given in Table `r bark_allometries`.

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
ss_table <- read.csv("tables_figures/sampling_details.csv", col.names = c("site", "species code", "n trees", "n cores", "n trees", "n cores", "sampled", "reconstructed*", "date range", "year effect analyzed"), stringsAsFactors = FALSE, check.names = FALSE)
kable(ss_table, format = "latex", booktabs = TRUE,   escape = F, linesep = c('\\addlinespace'), longtable = T)%>%
  add_header_above(c(" ", " ", "all"=2, "with DBH"=2, "DBH range" = 2, " ")) %>%
  column_spec(2, width = "1cm") %>%
  column_spec(3:6, width = "1cm") %>%
  column_spec(10, width = "1.5cm") %>%
  kable_styling(latex_options = c("scale_down", "hold_position","repeat_header"), protect_latex = T) 
```
*Maximum reconstructed diameters at breast height (DBH's) analyzed are less than maximum sampled DBH's because we discard size ranges with < 3 conspecific trees.

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
allometries_table <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO_dendro/data_processed/dbh_to_bark_allometries_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
kable(allometries_table, format = "latex", booktabs = TRUE,   escape = F, linesep = c('\\addlinespace'))%>%
  column_spec(1, italic = T) %>%
  column_spec(6, width = "3cm") %>%
  column_spec(7, width = "6cm") %>%
  row_spec(0, italic = F, bold = T) %>%
  kable_styling(latex_options = c("scale_down", "hold_position"), protect_latex = T)

# NOTE: references for this table are in the nocite list below table S5. Apparently refs in the first of two nocite sections don't appear in references list.
```
For assignments of species as proxies for those with out available bark allometries, see Table `r species_table`.

@@

PPT = precipitation, PET = potential evapotranspiration.

@@

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_SCBI_LITU_pre.png){ width=95% }

@@

![](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_SCBI_LITU_pet.png){ width=95%}

@@

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_Zofin_ABAL_wet.png){ width=95% }

@@

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_Zofin_ABAL_tmx.png){ width=95% }

@@

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_CedarBreaks_PSME_pre.png){ width=95% }

@@

![](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_CedarBreaks_PSME_tmx.png){ width=95% }

@@

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_ScottyCreek_PIMA_pre.png){ width=95% }

@@

![](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_ScottyCreek_PIMA_tmx.png){ width=95% }

@@

![`r GLS_BCNM_legend` Shown are responses for all three growth metrics examined here: ring width (RW), basal area increment (BAI), and aboveground biomass growth ($\Delta$AGB). `r GLS_generic_legend` PPT= precipitation.](tables_figures/SI_figures/composite_plots/BCI.png){ width=85% }

@@

![`r GLS_SCBI_legend` Shown are responses for all three growth metrics examined here: ring width (RW), basal area increment (BAI), and aboveground biomass growth ($\Delta$AGB). `r GLS_generic_legend`  PPT= precipitation, PET = potential evapotranspiration.](tables_figures/SI_figures/composite_plots/SCBI.png){ width=85% }

@@

![`r GLS_SCBI_year_legend` Shown are responses for all three growth metrics examined here: ring width (RW), basal area increment (BAI), and aboveground biomass growth ($\Delta$AGB). `r GLS_generic_legend` PPT= precipitation, PET = potential evapotranspiration.](tables_figures/SI_figures/composite_plots/SCBI_with_Year.png){ width=85% }

@@

![`r GLS_LT_year_legend` Shown are responses for all three growth metrics examined here: ring width (RW), basal area increment (BAI), and aboveground biomass growth ($\Delta$AGB). `r GLS_generic_legend` PPT= precipitation.](tables_figures/SI_figures/composite_plots/NewMexico_with_Year.png){ width=85% }

>>>>

# climwin plots  (turned off automatic transfer shortly prior to re-submission (8/31/21) because we had to manually edit plots)
#do.call(file.remove, list(list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-tree-rings/doc/manuscript/tables_figures/SI_figures/climwin_plots_combined/", full.names = TRUE))) # first remove old files
#flist <- list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-tree-rings/results/climwin_plots_combined/", full.names = TRUE)
#file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-tree-rings/doc/manuscript/tables_figures/SI_figures/climwin_plots_combined/")

@@

### Reconstructing diameter at breast height (DBH)

@@

$$DBH_Y  = DBH_{m} - 2*\left[r_{bark,m} - r_{bark,Y} + \sum_{y=Y}^{m} RW_{y} \right]$$
Here, $r_{bark}$ denotes bark thickness, which was estimated on a species-specific basis from allometries (detailed below; Tables `r species_table`, `r bark_allometries`).
If $DBH_m$ was taken before coring, RW values between $y=m$ and $y=Y$ were added instead of subtracted.
If $DBH_m$ was taken $\le$ 10 years after the last year of measurement in the core, we used the equation above, assuming that the $n$ missing measurements between the year of the last RW record in the core and year $m$ were equal to the last $n$ records in the core.

@@

$$DBH_Y  = 2*\left[r_{bark,Y} + \sum_{y=0}^{Y} RW_{y} \right]$$
Again, $r_{bark}$ was estimated on a species-specific basis from allometries (detailed below; Tables `r species_table`, `r bark_allometries`).

@@

### Estimating bark thickness
Regardless of the DBH reconstruction approach (outside-in or inside-out), accurate reconstructions require accounting for $r_{bark}$, particularly for species that can attain very thick bark (e.g., *Pseudotsuga menziesii*, *Pinus ponderosa* P. Lawson & C. Lawson).
For the majority of species, including all thick-barked species, we assigned allometries relating $r_{bark}$ to DBH, which were constructed for the species of interest or a related species (Tables `r species_table`, `r bark_allometries`).
When raw data on bark thickness data were available, we generated species-specific lognormal scaling relationships with the intercept forced to zero (Table `r bark_allometries`). 
When raw bark thickness data were not available, we applied published bark allometries from other sources (Table `r bark_allometries`).

@@

### Qualitative comparison

@@

### Quantitative comparison

@@

**Bark allometry field gives an ID number, corresponding to ID numbers in Table `r bark_allometries`, for the allometric equation used to estimate bark thickness. When neither raw data nor an allometric equation for the study species was available, we selected the most appropriate equation that could be located for similar species.

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
ss_table <- read.csv("tables_figures/sampling_details.csv", col.names = c("site", "species code", "n trees", "n cores", "n trees", "n cores", "sampled", "reconstructed*", "date range", "year effect analyzed"), stringsAsFactors = FALSE, check.names = FALSE)
kable(ss_table, format = "latex", booktabs = TRUE,   escape = F, linesep = c('\\addlinespace'), longtable = T)%>%
  add_header_above(c(" ", " ", "all"=2, "with DBH"=2, "DBH range" = 2, " ")) %>%
  column_spec(2, width = "1cm") %>%
  column_spec(3:6, width = "1cm") %>%
  column_spec(10, width = "1.5cm") %>%
  kable_styling(latex_options = c("scale_down", "hold_position","repeat_header"), protect_latex = T) 
```

Site and species codes are given in tables `r sites` and `r species_table`, respectively. Shown are numbers of trees and cores sampled and numbers for which DBH could be reconstructed (as described in Appendix `r `DBH_reconstruction`); ranges of tree sizes, in diameter at breast height (DBH; cm), as sampled and as reconstructed; date ranges analyzed; and whether or not data for the species met our criteria for analysis of year effects.

*Maximum reconstructed DBH's analyzed are less than maximum sampled DBH's because we discard size ranges with < 3 conspecific trees.

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
allometries_table <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO_dendro/data_processed/dbh_to_bark_allometries_table.csv",
                              col.names = c("ID", "species", "equation", "source type", "n", "DBH range (cm)", "sampling location", "source"),
                              stringsAsFactors = FALSE, check.names = FALSE)
kable(allometries_table, format = "latex", booktabs = TRUE,   escape = F, linesep = c('\\addlinespace'))%>%
  column_spec(2, width = "3cm") %>%
  column_spec(6, width = "2cm") %>%
  column_spec(7, width = "3cm") %>%
  column_spec(8, width = "4cm") %>%
  row_spec(0, italic = F, bold = T) %>%
  kable_styling(latex_options = c("scale_down", "hold_position"), protect_latex = T)

# NOTE: references for this table are in the nocite list below table S5. Apparently refs in the first of two nocite sections don't appear in references list.
```
Shown are allometric equations relating bark thickness (mm; *bark.mm*) to tree stem diameter at breast height (cm; *dbh.cm*). Equations were obtained from previous studies (source type = equation) or derived here from raw data (source type = data) as described in Appendix `r DBH_reconstruction`. For assignments of these allometries to the species included in this study, see Table `r species_table`.

@@

PPT = precipitation, PET = potential evapotranspiration. NA= not applicable (has not been analyzed).

@@

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_SCBI_LITU_pre.png){ width=90% }

@@

![](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_SCBI_LITU_pet.png){ width=90%}

@@

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_Zofin_ABAL_wet.png){ width=90% }

@@

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_Zofin_ABAL_tmx.png){ width=90% }

@@

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_CedarBreaks_PSME_pre.png){ width=90% }

@@

![](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_CedarBreaks_PSME_tmx.png){ width=90% }

@@

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_ScottyCreek_PIMA_pre.png){ width=90% }

@@

![](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_ScottyCreek_PIMA_tmx.png){ width=90% }

@@

![`r GLS_BCNM_legend` Shown are responses for all three growth metrics examined here: ring width (RW), basal area increment (BAI), and aboveground biomass growth ($\Delta$AGB). `r GLS_generic_legend` PPT= precipitation.](tables_figures/SI_figures/composite_plots/BCI.png){ width=80% }

@@

![`r GLS_SCBI_legend` Shown are responses for all three growth metrics examined here: ring width (RW), basal area increment (BAI), and aboveground biomass growth ($\Delta$AGB). `r GLS_generic_legend`  PPT= precipitation, PET = potential evapotranspiration.](tables_figures/SI_figures/composite_plots/SCBI.png){ width=80% }

@@

![`r GLS_SCBI_year_legend` Shown are responses for all three growth metrics examined here: ring width (RW), basal area increment (BAI), and aboveground biomass growth ($\Delta$AGB). `r GLS_generic_legend` PPT= precipitation, PET = potential evapotranspiration.](tables_figures/SI_figures/composite_plots/SCBI_with_Year.png){ width=80% }

@@

![`r GLS_LT_year_legend` Shown are responses for all three growth metrics examined here: ring width (RW), basal area increment (BAI), and aboveground biomass growth ($\Delta$AGB). `r GLS_generic_legend` PPT= precipitation.](tables_figures/SI_figures/composite_plots/NewMexico_with_Year.png){ width=80% }
